PaperOutline_ACEsCog

### 2022-01-26 PaperOutline_ACEsCog.Rmd
physical vs. emotional neglect
focused on analyses of physical vs emotional neglect instead of threat vs. deprviation

---
title: "SOBP_Analyses_BHRC_ABCD_ACEsCog"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(RColorBrewer)
```

## The goal of these analyses will be to compare ACEs (particularly Threat experiences) to cognition in two samples, the Brazilian High-Risk Cohort (BHRC) and the Adolescent Brain and Cognitive Development (ABCD) study.

@@

### Load and organize data:
```{r}
# Load and combine dataframes
sle_factors<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/thdp_20212904.rds")
ant_scores<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/ANT SCORES 100-1700ms RT cutoffs.csv")
demo_data<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/BHRCS_w0_Arielle.csv")
proposal.data<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/PhD_117BHRC_2021_10_02.rds") 
colnames(sle_factors)[2]<-"SUBJECTID"
colnames(demo_data)[2]<-"SUBJECTID"
combined.data1<-merge(demo_data,ant_scores,by="SUBJECTID")
proposal.data.w0<-proposal.data[proposal.data$redcap_event_name=="wave0_arm_1",c("ident", setdiff(colnames(proposal.data),colnames(combined.data1)))]
combined.data2<-merge(proposal.data.w0,sle_factors,by="ident")
all.data<-merge(combined.data1,combined.data2,by="SUBJECTID")
all.data$gender<-as.factor(all.data$gender)
all.data<-all.data[!is.na(all.data$gender),] #make sure gender variable is available for all participants

# Remove outliers (>3sd) on Alerting, Orienting, and Conflict scores
ALERT.3sd <- sd(all.data[!is.na(all.data$ALERT),]$ALERT)*3
ORIENT.3sd <- sd(all.data[!is.na(all.data$ORIENT),]$ORIENT)*3
CONFLICT.3sd <- sd(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)*3
ALERT.mean <- mean(all.data[!is.na(all.data$ALERT),]$ALERT)
ORIENT.mean <- mean(all.data[!is.na(all.data$ORIENT),]$ORIENT)
CONFLICT.mean <- mean(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)

# These will be separate data structures retaining as many subjects as possible for each of the 3 performance metrics
all.data.ALERT <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)),]
all.data.ORIENT <- all.data[(all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)),]
all.data.CONFLICT <- all.data[(all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]

# This is a complete set with subjects who have all 3 measures within the bounds of 3SD
all.data.complete.set <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)) & (all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)) & (all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]

# Add new data for BHRC
new_cog_data<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/BHRCS_Cog_IQ_TDE_w0.rds")
colnames(new_cog_data)[1]<-"SUBJECTID"
all.data.new<-merge(all.data.complete.set,new_cog_data[,c("SUBJECTID", setdiff(colnames(new_cog_data),colnames(all.data.complete.set)))],by="SUBJECTID")

@@

### Run linear models using Threat and Deprivation scores to predict cognition
# Note: Gender 1 = Male, Gender 2 = Female
# Covariates for age, gender, and random effects for school and state

# Significant:
```{r}
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ***
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***

summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # * (does not pass correction)
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***

summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ***
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***

p.adjust(c(0.000463,9.78e-07,0.0184,1.16e-07,0.000233,5.03e-06),method="bonferroni")

@@

# Plots:
```{r}
correlation <- cor.test(all.data.new$threat,all.data.new$WM)
ggplot(data=all.data.new,aes(threat,WM))+geom_point(size=1,colour='#70677D')+geom_smooth(method=lm,colour='black')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Working Memory') + xlab('Threat') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))

correlation <- cor.test(all.data.new$depriv,all.data.new$WM)
ggplot(data=all.data.new,aes(depriv,WM))+geom_point(size=1,colour='#7393C4')+geom_smooth(method=lm,colour='black')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Working Memory') + xlab('Deprivation') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
```

# Non-significant:
```{r}
summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns
summary(lmer(ORIENT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns
summary(lmer(CONFLICT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns

summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns
summary(lmer(ORIENT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns
summary(lmer(CONFLICT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns

summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+threat+depriv+threat*gender+depriv*gender,data=all.data.new)) # 0.055

summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ns
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ns
```

@@

### Load and organize data:
```{r}
nihtb<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/NihTB/abcd_tbss01.txt",header=TRUE)
nihtb.baseline<-nihtb[nihtb$eventname=="baseline_year_1_arm_1",]
lmt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/lmtp201.txt",header=TRUE)
lmt.baseline<-lmt[lmt$eventname=="baseline_year_1_arm_1",]
merged.data2 <- merge(nihtb.baseline,lmt.baseline[, c("subjectkey", setdiff(colnames(lmt.baseline),colnames(nihtb.baseline)))],by="subjectkey")
ravlt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/abcd_ps01.txt",header=TRUE)
ravlt.baseline<-ravlt[ravlt$eventname=="baseline_year_1_arm_1",]
merged.data3 <- merge(merged.data2,ravlt.baseline[, c("subjectkey", setdiff(colnames(ravlt.baseline),colnames(merged.data2)))],by="subjectkey")
trauma<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Strife/abcd_ptsd01.txt",header=TRUE)
trauma.baseline<-trauma[trauma$eventname=="baseline_year_1_arm_1",]
combined.data <- merge(merged.data3,trauma.baseline[, c("subjectkey", setdiff(colnames(trauma.baseline),colnames(merged.data3)))],by="subjectkey")
combined.data$nihtbx_flanker_uncorrected <- as.numeric(combined.data$nihtbx_flanker_uncorrected)
combined.data$nihtbx_picvocab_uncorrected <- as.numeric(combined.data$nihtbx_picvocab_uncorrected)
combined.data$nihtbx_list_uncorrected <- as.numeric(combined.data$nihtbx_list_uncorrected)
combined.data$nihtbx_pattern_uncorrected <- as.numeric(combined.data$nihtbx_pattern_uncorrected)
combined.data$nihtbx_cardsort_uncorrected <- as.numeric(combined.data$nihtbx_cardsort_uncorrected)
combined.data$nihtbx_picture_uncorrected <- as.numeric(combined.data$nihtbx_picture_uncorrected)
combined.data$nihtbx_reading_uncorrected <- as.numeric(combined.data$nihtbx_reading_uncorrected)
combined.data$lmt_scr_perc_correct <- as.numeric(combined.data$lmt_scr_perc_correct)
combined.data$pea_ravlt_sd_trial_i_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_i_tc)
combined.data$pea_ravlt_sd_trial_ii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_ii_tc)
combined.data$pea_ravlt_sd_trial_iii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iii_tc)
combined.data$pea_ravlt_sd_trial_iv_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iv_tc)
combined.data$pea_ravlt_sd_trial_v_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_v_tc)
combined.data$interview_age <- as.numeric(combined.data$interview_age)

@@

abcd.data<-merge(abcd.data,neglect.data[,c("subjectkey", setdiff(colnames(neglect.data),colnames(abcd.data)))],by="subjectkey")

# Calculate subscales of deprivation (emotional and physical neglect) - see this ref for motivation https://www.sciencedirect.com/science/article/pii/S2352289518300821
abcd.data$acceptance <- rowSums(data.frame(as.numeric(abcd.data$crpbi_parent1_y),as.numeric(abcd.data$crpbi_parent2_y),as.numeric(abcd.data$crpbi_parent3_y),as.numeric(abcd.data$crpbi_parent4_y),as.numeric(abcd.data$crpbi_parent5_y)))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
```

# The goal of these analyses will be to compare ACEs (particularly Threat experiences) to cognition in two samples, the Brazilian High-Risk Cohort (BHRC) and the Adolescent Brain and Cognitive Development (ABCD) study.

@@

### Load and organize data:
```{r}
# Load and combine dataframes
sle_factors<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/thdp_20212904.rds")
#ant_scores<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/ANT SCORES 100-1700ms RT cutoffs.csv")
demo_data<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/BHRCS_w0_Arielle.csv")
proposal.data<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/PhD_117BHRC_2021_10_02.rds") 
data_depriv_emoneg<-readRDS("/Users/ariellekeller/Downloads/Arielle_BHRC_2021_12_26_2.rds")
sle_and_depriv <- merge(sle_factors[,c(1,2,4,5)],data_depriv_emoneg,by="ident") #don't duplicate redcap event name
colnames(sle_and_depriv)[2]<-"SUBJECTID"
colnames(demo_data)[2]<-"SUBJECTID"
#combined.data1<-merge(demo_data,ant_scores,by="SUBJECTID")
proposal.data.w0<-proposal.data[proposal.data$redcap_event_name=="wave0_arm_1",c("ident", setdiff(colnames(proposal.data),colnames(demo_data)))]
combined.data2<-merge(proposal.data.w0,sle_and_depriv[,c("ident", setdiff(colnames(sle_and_depriv),colnames(proposal.data.w0)))],by="ident")
all.data<-merge(demo_data,combined.data2[,c("SUBJECTID", setdiff(colnames(combined.data2),colnames(demo_data)))],by="SUBJECTID")
all.data$gender<-as.factor(all.data$gender)
all.data<-all.data[!is.na(all.data$gender),] #make sure gender variable is available for all participants
# 
# # Remove outliers (>3sd) on Alerting, Orienting, and Conflict scores
# ALERT.3sd <- sd(all.data[!is.na(all.data$ALERT),]$ALERT)*3
# ORIENT.3sd <- sd(all.data[!is.na(all.data$ORIENT),]$ORIENT)*3
# CONFLICT.3sd <- sd(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)*3
# ALERT.mean <- mean(all.data[!is.na(all.data$ALERT),]$ALERT)
# ORIENT.mean <- mean(all.data[!is.na(all.data$ORIENT),]$ORIENT)
# CONFLICT.mean <- mean(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)
# 
# # These will be separate data structures retaining as many subjects as possible for each of the 3 performance metrics
# all.data.ALERT <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)),]
# all.data.ORIENT <- all.data[(all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)),]
# all.data.CONFLICT <- all.data[(all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]
# 
# # This is a complete set with subjects who have all 3 measures within the bounds of 3SD
# all.data.complete.set <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)) & (all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)) & (all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]

# Add new data for BHRC
new_cog_data<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/BHRCS_Cog_IQ_TDE_w0.rds")
colnames(new_cog_data)[1]<-"SUBJECTID"
all.data.new<-merge(all.data,new_cog_data[,c("SUBJECTID", setdiff(colnames(new_cog_data),colnames(all.data)))],by="SUBJECTID")

@@

```{r}
# Remove outliers (>3sd)
EF.3sd <- sd(all.data.new[!is.na(all.data.new$EF),]$EF)*3
WM.3sd <- sd(all.data.new[!is.na(all.data.new$WM),]$WM)*3
TP.3sd <- sd(all.data.new[!is.na(all.data.new$TP),]$TP)*3
IC.3sd <- sd(all.data.new[!is.na(all.data.new$IC),]$IC)*3
EF.mean <- mean(all.data.new[!is.na(all.data.new$EF),]$EF)
WM.mean <- mean(all.data.new[!is.na(all.data.new$WM),]$WM)
TP.mean <- mean(all.data.new[!is.na(all.data.new$TP),]$TP)
IC.mean <- mean(all.data.new[!is.na(all.data.new$IC),]$IC)

# This is a complete set with subjects who have all 4 cog measures within the bounds of 3SD
all.data.complete <- all.data.new[(all.data.new$EF>(EF.mean-EF.3sd) & all.data.new$EF<(EF.mean+EF.3sd)) & (all.data.new$WM>(WM.mean-WM.3sd) & all.data.new$WM<(WM.mean+WM.3sd)) & (all.data.new$IC>(IC.mean-IC.3sd) & all.data.new$IC<(IC.mean+IC.3sd)) & (all.data.new$TP>(TP.mean-TP.3sd) & all.data.new$TP<(TP.mean+TP.3sd)),]

@@

## Run linear models using Threat and Deprivation scores to predict cognition
### Note: Gender 1 = Male, Gender 2 = Female
### Covariates for age, gender, and random effects for school and state

### Significant:
```{r}
# summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ***
# summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***
# 
# summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # * (does not pass correction)
# summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***
# 
# summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ***
# summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ***
# 
# p.adjust(c(0.000463,9.78e-07,0.0184,1.16e-07,0.000233,5.03e-06),method="bonferroni")

### Plots:
```{r}
# correlation <- cor.test(all.data.new$threat,all.data.new$WM)
# ggplot(data=all.data.new,aes(threat,WM))+geom_point(size=1,colour='#70677D')+geom_smooth(method=lm,colour='black')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
# panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Working Memory') + xlab('Threat') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
# 
# correlation <- cor.test(all.data.new$depriv,all.data.new$WM)
# ggplot(data=all.data.new,aes(depriv,WM))+geom_point(size=1,colour='#7393C4')+geom_smooth(method=lm,colour='black')+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
# panel.background = element_blank(), axis.line = element_line(colour = "black")) +ylab('Working Memory') + xlab('Deprivation') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
```

### For the effects where there was also a significant effect of gender, look for an interaction (there are none)
```{r}
# summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+threat+threat*gender,data=all.data.new)) # ns
# summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+depriv+depriv*gender,data=all.data.new)) # ns
# 
# summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+threat+threat*gender,data=all.data.new)) # ns
# summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+depriv+depriv*gender,data=all.data.new)) # ns
```

### Look for specificity on the type of ACE
#### Note that the correlation between threat and deprivation here is only 0.386
```{r}
# cor.test(all.data.new$threat,all.data.new$depriv)
# summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+threat+depriv,data=all.data.new)) # specific to depriv ***
# summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+threat+depriv,data=all.data.new)) # specific to depriv ***
# summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+threat+depriv,data=all.data.new)) # both threat * and depriv ***
```

## Create a correlation matrix of all the important variables
### Since the cognitive variables are more highly correlated with each other, we won't test specificity in a single model
```{r}
# important.variables <- all.data.new[,c("age","gender","threat","depriv","EF","WM","TP","IC")]
# important.variables$gender <- as.numeric(important.variables$gender)
# colnames(important.variables)<-c("Age","Gender","Threat","Depriv","Exec. Fxn","Working Mem","Temp. Proc.","Inhib. Cont.")
# corrplot(cor(important.variables,use="complete.obs"),method=c("number"), tl.col = "black", col=brewer.pal(n=8, name="Spectral"))
```

### Non-significant:
```{r}
# summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns
# summary(lmer(ORIENT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns
# summary(lmer(CONFLICT~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) #ns
# 
# summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns
# summary(lmer(ORIENT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns
# summary(lmer(CONFLICT~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) #ns
# 
# summary(lmer(ALERT~age+gender+(1|state)+(1|schoolID)+threat+depriv+threat*gender+depriv*gender,data=all.data.new)) # 0.055
# 
# summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+threat,data=all.data.new)) # ns
# summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+depriv,data=all.data.new)) # ns
```

### Look for specificity on Physical vs Emotional Neglect (as an replication of ABCD findings)
```{r}
# Calculate emotional neglect variable by averaging across relevant items (the "depriv" one will represent physical neglect)
all.data.complete<-all.data.complete[all.data.complete$fat38a<50,]
all.data.complete$emo_neglect <- -rowMeans(data.frame(as.numeric(all.data.complete$fat38a)/3,as.numeric(all.data.complete$fat39a)/3),na.rm=TRUE) #,as.numeric(all.data.complete$fes18),as.numeric(all.data.complete$fes1),as.numeric(all.data.complete$fes42),as.numeric(all.data.complete$fes48))

# Verify that correlation is low: it's 0.222
cor.test(all.data.complete$depriv,all.data.complete$emo_neglect)

# Test specificity
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete)) # both sig
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete)) # both sig
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete)) # both sig 
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete)) # just depriv

p.adjust(c(0.005610,1.72e-09,0.0030,3.22e-08,0.0111,5.23e-09,0.323230,0.014438),method="bonferroni")

# try restricting age to ABCD range:
all.data.complete.ageres <- all.data.complete[all.data.complete$age<11.1,]
all.data.complete.ageres <- all.data.complete.ageres[all.data.complete.ageres$age>8.9,]
all.data.complete.ageres<-all.data.complete.ageres[all.data.complete.ageres$fat38a<50,]
all.data.complete.ageres$emo_neglect <- -rowMeans(data.frame(as.numeric(all.data.complete.ageres$fat38a)/3,as.numeric(all.data.complete.ageres$fat39a)/3),na.rm=TRUE) 
cor.test(all.data.complete.ageres$depriv,all.data.complete.ageres$emo_neglect)

# Test specificity
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete.ageres)) # both sig
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete.ageres)) # both sig
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete.ageres)) # both sig 
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+depriv,data=all.data.complete.ageres)) # just depriv

p.adjust(c(0.0202,3.91e-05,0.0556,1.12e-06,0.020554,0.000262,0.4224,0.0588),method="bonferroni") 
```

## Create a correlation matrix of all the important variables
### Since the cognitive variables are more highly correlated with each other, we won't test specificity in a single model
```{r}
important.variables <- all.data.complete.ageres[,c("age","gender","emo_neglect","depriv","EF","WM","TP","IC")]
important.variables$gender <- as.numeric(important.variables$gender)
colnames(important.variables)<-c("Age","Gender","EmoNeglect","PhysNeglect","Exec. Fxn","Working Mem","Temp. Proc.","Inhib. Cont.")
corrplot(cor(important.variables,use="complete.obs"),method=c("number"), tl.col = "black", col=brewer.pal(n=8, name="Spectral"))
```

## Plots
```{r}
cols <- brewer.pal(10,"Paired")

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_PhysNeg_TP.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=depriv,y=TP)) +
  geom_point(colour=cols[2],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Physical Neglect") + ylab("Temporal Processing") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_PhysNeg_EF.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=depriv,y=EF)) +
  geom_point(colour=cols[4],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Physical Neglect") + ylab("Exec. Fxn.") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_PhysNeg_WM.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=depriv,y=WM)) +
  geom_point(colour=cols[8],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Physical Neglect") + ylab("Working Memory") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_PhysNeg_IC.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=depriv,y=IC)) +
  geom_point(colour=cols[10],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Physical Neglect") + ylab("Inhibitory Control") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_EmoNeg_TP.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=emo_neglect,y=TP)) +
  geom_point(colour=cols[2],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Emotional Neglect") + ylab("Temporal Processing") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_EmoNeg_EF.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=emo_neglect,y=EF)) +
  geom_point(colour=cols[4],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Emotional Neglect") + ylab("Exec. Fxn.") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_EmoNeg_WM.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=emo_neglect,y=WM)) +
  geom_point(colour=cols[8],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Emotional Neglect") + ylab("Working Memory") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()

 png("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/PhysEmoNeglect_ABCD_BHRC/Scatterplots/BHRC_EmoNeg_IC.png",
     width = 3000, 
     height = 3000,
     res = 850)
ggplot(all.data.complete.ageres, aes(x=emo_neglect,y=IC)) +
  geom_point(colour=cols[10],size=1) +
  geom_smooth(method=lm,color="black") + xlab("Emotional Neglect") + ylab("Inhibitory Control") + theme(legend.position = "none") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black"))
dev.off()
```

@@

### Load and organize data:
```{r}
nihtb<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/NihTB/abcd_tbss01.txt",header=TRUE)
nihtb.baseline<-nihtb[nihtb$eventname=="baseline_year_1_arm_1",]
lmt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/lmtp201.txt",header=TRUE)
lmt.baseline<-lmt[lmt$eventname=="baseline_year_1_arm_1",]
merged.data2 <- merge(nihtb.baseline,lmt.baseline[, c("subjectkey", setdiff(colnames(lmt.baseline),colnames(nihtb.baseline)))],by="subjectkey")
ravlt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/abcd_ps01.txt",header=TRUE)
ravlt.baseline<-ravlt[ravlt$eventname=="baseline_year_1_arm_1",]
combined.data <- merge(merged.data2,ravlt.baseline[, c("subjectkey", setdiff(colnames(ravlt.baseline),colnames(merged.data2)))],by="subjectkey")
# trauma<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Strife/abcd_ptsd01.txt",header=TRUE)
# trauma.baseline<-trauma[trauma$eventname=="baseline_year_1_arm_1",]
# combined.data <- merge(merged.data3,trauma.baseline[, c("subjectkey", setdiff(colnames(trauma.baseline),colnames(merged.data3)))],by="subjectkey")
combined.data$nihtbx_flanker_uncorrected <- as.numeric(combined.data$nihtbx_flanker_uncorrected)
combined.data$nihtbx_picvocab_uncorrected <- as.numeric(combined.data$nihtbx_picvocab_uncorrected)
combined.data$nihtbx_list_uncorrected <- as.numeric(combined.data$nihtbx_list_uncorrected)
combined.data$nihtbx_pattern_uncorrected <- as.numeric(combined.data$nihtbx_pattern_uncorrected)
combined.data$nihtbx_cardsort_uncorrected <- as.numeric(combined.data$nihtbx_cardsort_uncorrected)
combined.data$nihtbx_picture_uncorrected <- as.numeric(combined.data$nihtbx_picture_uncorrected)
combined.data$nihtbx_reading_uncorrected <- as.numeric(combined.data$nihtbx_reading_uncorrected)
combined.data$lmt_scr_perc_correct <- as.numeric(combined.data$lmt_scr_perc_correct)
combined.data$pea_ravlt_sd_trial_i_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_i_tc)
combined.data$pea_ravlt_sd_trial_ii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_ii_tc)
combined.data$pea_ravlt_sd_trial_iii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iii_tc)
combined.data$pea_ravlt_sd_trial_iv_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iv_tc)
combined.data$pea_ravlt_sd_trial_v_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_v_tc)
combined.data$interview_age <- as.numeric(combined.data$interview_age)

@@

abcd.data<-merge(abcd.data.almost,neglect.data[,c("subjectkey", setdiff(colnames(neglect.data),colnames(abcd.data.almost)))],by="subjectkey")

# Calculate subscales of deprivation (emotional and physical neglect) - see this ref for motivation https://www.sciencedirect.com/science/article/pii/S2352289518300821
abcd.data$acceptance <- rowMeans(data.frame(as.numeric(abcd.data$crpbi_parent1_y),as.numeric(abcd.data$crpbi_parent2_y),as.numeric(abcd.data$crpbi_parent3_y),as.numeric(abcd.data$crpbi_parent4_y),as.numeric(abcd.data$crpbi_parent5_y)))

abcd.data$monitoring <- rowMeans(data.frame(as.numeric(abcd.data$parent_monitor_q1_y),as.numeric(abcd.data$parent_monitor_q2_y),as.numeric(abcd.data$parent_monitor_q3_y),as.numeric(abcd.data$parent_monitor_q4_y),as.numeric(abcd.data$parent_monitor_q5_y)))
